---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.path = "man/figures/README-",
  fig.align = "center",
  out.width = "100%",
  results = "hold",
  comment = "#>"
)
```

# alpaca

<!-- badges: start -->
[![R-CMD-check](https://github.com/dereckscompany/alpaca/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dereckscompany/alpaca/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

An R API wrapper for the [Alpaca](https://alpaca.markets/) trading platform. Provides `R6` classes for market data, stock trading, options, account management, and positions. Supports both synchronous and asynchronous (promise based) operation via `httr2`.

## Disclaimer

This software is provided "as is", without warranty of any kind. **This package interacts with live brokerage accounts and can execute real trades involving real money.** By using this package you accept full responsibility for any financial losses, erroneous transactions, or other damages that may result. Always test with paper trading first, use API key permissions to restrict access to only what you need, and never share your API credentials. The author(s) and contributor(s) are not liable for any financial loss or damage arising from the use of this software.

We invite you to read the source code and make contributions if you find a bug or wish to make an improvement.

## Design Philosophy

All API responses are returned as `data.table` objects with three transformations applied:

1. **snake_case column names** -- camelCase keys from the JSON response (e.g. `latestTrade`, `nextClose`) are converted to snake_case (`latest_trade`, `next_close`). A handful of endpoints additionally expand abbreviated single-letter keys returned by Alpaca's bar/trade/quote feeds (e.g. `t` → `timestamp`, `p` → `price`, `o`/`h`/`l`/`c` → `open`/`high`/`low`/`close`, `vw` → `volume_weighted_avg_price`) so the resulting column names are self-describing. The full mapping per endpoint is documented in the method-level `@return` blocks.

2. **Timestamps to POSIXct where the field is clearly a market-data timestamp** -- Numeric epoch fields and obvious RFC-3339 fields on market-data endpoints (bars, trades, quotes, snapshots, orderbook) are converted to `POSIXct` under their snake_case name. Some textual `created_at` / `updated_at` fields on lower-level endpoints (notably news articles) are left as `character` strings; see the relevant `@return` for the exact type per column.

3. **One entity = one row, no list columns** -- For every endpoint normalised under the shape policy (see *Data-shape conventions* below) the returned `data.table` contains no list columns. A small number of legacy endpoints that go through the generic parser may still surface nested arrays as list columns; those are called out individually in their `@return`.

**No fields are dropped** by the normalising parsers (`parse_trades`, `parse_quotes`, `parse_news`, `parse_snapshot`, `parse_asset`/`parse_assets`, `parse_order`/`parse_orders`, `parse_watchlist`). Every scalar and array field present in the Alpaca response surfaces as a column on the resulting `data.table` -- collapsed, exploded, or wide-prefixed per the shape policy. If you don't need a column, drop it yourself.

## Data-shape conventions

Alpaca's JSON responses sometimes contain nested arrays and objects. To keep returns intuitive (and consistent across the `alpaca` / `binance` / `kucoin` packages), every method follows one rule: **identify the entity for the endpoint, and return one row per entity**. The four cases:

| Nested shape | Treatment | Example |
| --- | --- | --- |
| Array of plain strings (`conditions`, `attributes`, `permissions`, `symbols`, `image_urls`) | Collapsed into a single character column joined by `;`. | `dt$attributes` -> `"fractional_eh_enabled;has_options;overnight_tradable"` |
| Array of objects (orderbook levels, watchlist assets, multi-leg orders) | Exploded to long format with parent fields replicated. A position-index column is added when order matters (`level`, `leg_index`). | `get_crypto_orderbook("BTC/USD")` -> one row per `(side, level)`. |
| Fixed-schema nested object (snapshot bars, account configurations) | Flattened to wide `parent_child` columns. | `get_account()` -> `admin_configurations_max_options_trading_level`. |
| Empty / null array | `NA_character_` (no list cells). Empty responses -> empty `data.table` (not stub rows). | An asset with no attributes -> `attributes = NA`. |

### Recovering the original values

Filter with `grepl`:

```r
# All assets that support options trading
options_assets <- assets[grepl("has_options", attributes)]

# All trades reported on the consolidated tape
trades_on_tape <- trades[grepl("T", conditions)]
```

Split back into the original character vector:

```r
strsplit(dt$attributes[1], ";", fixed = TRUE)[[1]]
# [1] "fractional_eh_enabled" "has_options" "overnight_tradable"
```

URL-encoded fields (currently just news `image_urls`, where Alpaca legitimately uses `;` inside some URLs) round-trip via `URLdecode()`. The encoding is **lossless** in both directions: each URL is double-encoded (`%` → `%25`, then `;` → `%3B`) before joining, so a single `URLdecode()` after splitting recovers the original string even when the upstream URL already contained `%3B` or other percent-escapes:

```r
urls <- strsplit(news$image_urls[1], ";", fixed = TRUE)[[1]]
urls <- vapply(urls, URLdecode, character(1))
```

For parallel arrays like `image_sizes` / `image_urls`, when one element omits a per-image field the missing value becomes an **empty token** (e.g. `"large;"`), never the literal string `"NA"` — so a real `"NA"` value remains unambiguous from a missing one.

### Multi-leg orders (`bracket` / `oco` / `oto`)

Multi-leg orders return a flat `data.table` with one row per "order" (parent and legs treated equally). Two helper columns disambiguate:

* `leg_index` (integer) -- `NA` on the parent row; `1, 2, ...` on each leg.
* `parent_order_id` (character) -- `NA` on the parent row; the parent's `id` on each leg.

```r
# Just the parent orders (filter out legs)
dt[is.na(parent_order_id)]

# All legs of one specific bracket
dt[parent_order_id == "<parent-uuid>"]
```

## Installation

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("dereckscompany/alpaca")
```

## Setup

```{r}
# special mock for local build
box::use(
  alpaca[
    get_api_keys,
    get_base_url,
    get_data_base_url
  ],
  ./tests/testthat/mock_router[mock_router]
)

KEYS <- get_api_keys(
  api_key = "fake-key",
  api_secret = "fake-secret"
)

TBASE <- "https://paper-api.alpaca.markets"
DBASE <- "https://data.alpaca.markets"

options(httr2_mock = mock_router)

# normal imports
box::use(
  alpaca[
    AlpacaMarketData,
    AlpacaTrading,
    AlpacaAccount,
    AlpacaOptions
  ]
)
```

Set your API credentials as environment variables in `.Renviron`:

```bash
ALPACA_API_KEY = your-api-key
ALPACA_API_SECRET = your-api-secret
ALPACA_API_ENDPOINT = https://paper-api.alpaca.markets
```

If you don't have a key, visit the [Alpaca dashboard](https://app.alpaca.markets/).

## Quick Start -- Market Data

All endpoints require authentication (even market data).

```{r}
market <- AlpacaMarketData$new(keys = KEYS, base_url = TBASE, data_base_url = DBASE)
```

### Historical Bars

```{r}
bars <- market$get_bars(
  symbol = "AAPL",
  timeframe = "1Day",
  start = "2024-01-01",
  end = "2024-01-31"
)
bars[]
```

### Latest Trade

```{r}
trade <- market$get_latest_trade(symbol = "AAPL")
trade[]
```

### Latest Quote (NBBO)

```{r}
quote <- market$get_latest_quote(symbol = "AAPL")
quote[]
```

### Market Clock

```{r}
market$get_clock()
```

### Available Assets

```{r}
assets <- market$get_assets(status = "active", asset_class = "us_equity")
assets[]
```

### Market News

```{r}
news <- market$get_news(symbols = "AAPL", limit = 5)
news[, .(headline, source, created_at)]
```

## Account

```{r}
acct <- AlpacaAccount$new(keys = KEYS, base_url = TBASE)
```

### Account Info

```{r}
info <- acct$get_account()
info[, .(status, equity, buying_power, cash)]
```

### Open Positions

```{r}
acct$get_positions()
```

### Portfolio History

```{r}
acct$get_portfolio_history(period = "1M", timeframe = "1D")
```

## Trading

```{r}
trading <- AlpacaTrading$new(keys = KEYS, base_url = TBASE)
```

### Place a Limit Order

```{r}
order <- trading$add_order(
  symbol = "AAPL",
  side = "buy",
  type = "limit",
  time_in_force = "day",
  qty = 1,
  limit_price = 150
)
order[, .(id, symbol, side, type, status, limit_price)]
```

### List Open Orders

```{r}
trading$get_orders(status = "open")
```

### Cancel an Order

```{r}
trading$cancel_order(order_id = "order-uuid-123")
```

## Options

```{r}
opts <- AlpacaOptions$new(keys = KEYS, base_url = TBASE, data_base_url = DBASE)
```

### List Contracts

```{r}
contracts <- opts$get_contracts(
  underlying_symbols = "AAPL",
  type = "call",
  expiration_date_gte = "2024-06-01",
  limit = 10
)
contracts[, .(symbol, type, strike_price, expiration_date)]
```

### Options Chain

```{r}
chain <- opts$get_option_chain(underlying_symbol = "AAPL", type = "call")
chain[]
```

## Short Selling

Alpaca integrates margin and short selling directly into the standard trading
API -- no special classes needed. To short a stock, simply place a sell order
for a symbol you don't own:

```{r, eval = FALSE}
# Short 100 shares of AAPL at market
trading$add_order(
  symbol = "AAPL",
  side = "sell",
  type = "market",
  time_in_force = "day",
  qty = 100
)

# Close the short (buy to cover)
acct$close_position(symbol_or_id = "AAPL")
```

## Bulk Historical Data

The `alpaca_backfill_bars()` function downloads historical bar data for
multiple symbols and timeframes with CSV-based resume support:

```{r, eval = FALSE}
dt <- alpaca_backfill_bars(
  symbols = c("AAPL", "MSFT", "TSLA"),
  timeframes = c("1Day", "1Hour"),
  start = "2020-01-01",
  path = "data/bars.csv"
)

# Resume an interrupted download -- already-completed combos are skipped
dt <- alpaca_backfill_bars(
  symbols = c("AAPL", "MSFT", "TSLA", "GOOGL"),
  timeframes = c("1Day", "1Hour"),
  start = "2020-01-01",
  path = "data/bars.csv"
)
```

## Sample Data

The package includes a bundled dataset of simulated AAPL daily bars for
testing and demonstration:

```{r}
data(alpaca_aapl_1day_bars, package = "alpaca")
head(alpaca_aapl_1day_bars)
```

## Async Usage

All classes accept `async = TRUE`, causing methods to return promises.
Use `coro::async()` to write sequential-looking async code:

```{r}
market_async <- AlpacaMarketData$new(async = TRUE)

main <- coro::async(function() {
  bars <- await(market_async$get_bars(
    symbol = "AAPL",
    timeframe = "1Day",
    start = "2024-01-01",
    end = "2024-01-31"
  ))
  clock <- await(market_async$get_clock())

  print(bars)
  cat("Market open:", clock$is_open, "\n")
})

main()
while (!later::loop_empty()) {
  later::run_now()
}
```

## Available Classes

| Class | Purpose |
|-------|---------|
| `AlpacaMarketData` | Historical bars, latest trades/quotes, snapshots, assets, calendar, clock, news, screener, corporate actions |
| `AlpacaTrading` | Place, modify, cancel, and query orders |
| `AlpacaAccount` | Account info, positions, portfolio history, activities, watchlists |
| `AlpacaOptions` | Options contracts, bars, trades, quotes, snapshots, chain |

Standalone function: `alpaca_backfill_bars()` -- bulk historical bar download with CSV resume.

## Citation

If you use this package in your work, please cite it:

```r
citation("alpaca")
```

> Mezquita, D. (2026). alpaca: R API Wrapper to Alpaca Trading Platform. R package version 0.1.0.

## Licence

MIT &copy; [Dereck Mezquita](https://github.com/dereckmezquita) [![ORCID](https://img.shields.io/badge/ORCID-0000--0002--9307--6762-green)](https://orcid.org/0000-0002-9307-6762). See [LICENSE.md](LICENSE.md) for the full text, including the citation clause.
