---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.path = "man/figures/README-",
  fig.align = "center",
  out.width = "100%",
  collapse = TRUE,
  comment = "#>"
)

box::use(
  alpaca[
    AlpacaMarketData, AlpacaTrading, AlpacaAccount, AlpacaOptions,
    get_api_keys, get_base_url, get_data_base_url,
    alpaca_backfill_bars
  ]
)

box::use(./tests/testthat/mock_router[mock_router])

KEYS <- get_api_keys(
  api_key = "fake-key",
  api_secret = "fake-secret"
)
TBASE <- "https://paper-api.alpaca.markets"
DBASE <- "https://data.alpaca.markets"

options(httr2_mock = mock_router)
```

# alpaca

<!-- badges: start -->
[![R-CMD-check](https://github.com/dereckmezquita/alpaca/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dereckmezquita/alpaca/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

An R API wrapper for the [Alpaca](https://alpaca.markets/) trading platform. Provides R6 classes for market data, stock trading, options, account management, and positions. Supports both synchronous and asynchronous (promise-based) operation via httr2.

## Disclaimer

This software is provided "as is", without warranty of any kind. **This package interacts with live brokerage accounts and can execute real trades involving real money.** By using this package you accept full responsibility for any financial losses, erroneous transactions, or other damages that may result. Always test with paper trading first, use API key permissions to restrict access to only what you need, and never share your API credentials. The author(s) and contributor(s) are not liable for any financial loss or damage arising from the use of this software.

We invite you to read the source code and make contributions if you find a bug or wish to make an improvement.

## Design Philosophy

All API responses are returned as `data.table` objects with two transformations applied:

1. **snake_case column names** -- camelCase keys from the JSON response (e.g. `latestTrade`, `nextClose`) are converted to snake_case (`latest_trade`, `next_close`) via a mechanical transformation. No columns are renamed beyond this.

2. **Timestamps to POSIXct** -- Columns containing RFC-3339 timestamps or epoch seconds are converted to `POSIXct` in-place under their snake_case name.

That's it. **No fields are dropped and no columns are renamed** beyond the camelCase-to-snake_case conversion. If a column exists in the Alpaca API response, it will exist in the returned `data.table`. If you don't need a column, drop it yourself.

## Installation

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("dereckmezquita/alpaca")
```

## Setup

Set your API credentials as environment variables in `.Renviron`:

```bash
ALPACA_API_KEY = your-api-key
ALPACA_API_SECRET = your-api-secret
ALPACA_API_ENDPOINT = https://paper-api.alpaca.markets
```

If you don't have a key, visit the [Alpaca dashboard](https://app.alpaca.markets/).

## Quick Start -- Market Data

All endpoints require authentication (even market data).

```{r}
market <- AlpacaMarketData$new(keys = KEYS, base_url = TBASE, data_base_url = DBASE)
```

```{r echo = FALSE}
market <- AlpacaMarketData$new(keys = KEYS, base_url = TBASE, data_base_url = DBASE)
```

### Historical Bars

```{r}
bars <- market$get_bars("AAPL", timeframe = "1Day", start = "2024-01-01", end = "2024-01-31")
bars[]
```

### Latest Trade

```{r}
trade <- market$get_latest_trade("AAPL")
trade[]
```

### Latest Quote (NBBO)

```{r}
quote <- market$get_latest_quote("AAPL")
quote[]
```

### Market Clock

```{r}
market$get_clock()
```

### Available Assets

```{r}
assets <- market$get_assets(status = "active", asset_class = "us_equity")
assets[]
```

### Market News

```{r}
news <- market$get_news(symbols = "AAPL", limit = 5)
news[, .(headline, source, created_at)]
```

## Account

```{r}
acct <- AlpacaAccount$new(keys = KEYS, base_url = TBASE)
```

```{r echo = FALSE}
acct <- AlpacaAccount$new(keys = KEYS, base_url = TBASE)
```

### Account Info

```{r}
info <- acct$get_account()
info[, .(status, equity, buying_power, cash)]
```

### Open Positions

```{r}
acct$get_positions()
```

### Portfolio History

```{r}
acct$get_portfolio_history(period = "1M", timeframe = "1D")
```

## Trading

```{r}
trading <- AlpacaTrading$new(keys = KEYS, base_url = TBASE)
```

```{r echo = FALSE}
trading <- AlpacaTrading$new(keys = KEYS, base_url = TBASE)
```

### Place a Limit Order

```{r}
order <- trading$add_order(
  symbol = "AAPL", side = "buy", type = "limit",
  time_in_force = "day", qty = 1, limit_price = 150
)
order[, .(id, symbol, side, type, status, limit_price)]
```

### List Open Orders

```{r}
trading$get_orders(status = "open")
```

### Cancel an Order

```{r}
trading$cancel_order("order-uuid-123")
```

## Options

```{r}
opts <- AlpacaOptions$new(keys = KEYS, base_url = TBASE, data_base_url = DBASE)
```

```{r echo = FALSE}
opts <- AlpacaOptions$new(keys = KEYS, base_url = TBASE, data_base_url = DBASE)
```

### List Contracts

```{r}
contracts <- opts$get_contracts(
  underlying_symbols = "AAPL",
  type = "call",
  expiration_date_gte = "2024-06-01",
  limit = 10
)
contracts[, .(symbol, type, strike_price, expiration_date)]
```

### Options Chain

```{r}
chain <- opts$get_option_chain("AAPL", type = "call")
chain[]
```

## Short Selling

Alpaca integrates margin and short selling directly into the standard trading
API -- no special classes needed. To short a stock, simply place a sell order
for a symbol you don't own:

```{r, eval = FALSE}
# Short 100 shares of AAPL at market
trading$add_order(
  symbol = "AAPL", side = "sell", type = "market",
  time_in_force = "day", qty = 100
)

# Close the short (buy to cover)
acct$close_position("AAPL")
```

## Bulk Historical Data

The `alpaca_backfill_bars()` function downloads historical bar data for
multiple symbols and timeframes with CSV-based resume support:

```{r, eval = FALSE}
dt <- alpaca_backfill_bars(
  symbols = c("AAPL", "MSFT", "TSLA"),
  timeframes = c("1Day", "1Hour"),
  start = "2020-01-01",
  path = "data/bars.csv"
)

# Resume an interrupted download -- already-completed combos are skipped
dt <- alpaca_backfill_bars(
  symbols = c("AAPL", "MSFT", "TSLA", "GOOGL"),
  timeframes = c("1Day", "1Hour"),
  start = "2020-01-01",
  path = "data/bars.csv"
)
```

## Sample Data

The package includes a bundled dataset of simulated AAPL daily bars for
testing and demonstration:

```{r}
data(alpaca_aapl_1day_bars, package = "alpaca")
head(alpaca_aapl_1day_bars)
```

## Async Usage

All classes accept `async = TRUE`, causing methods to return promises.
Use `coro::async()` to write sequential-looking async code:

```{r, eval = FALSE}
market_async <- AlpacaMarketData$new(async = TRUE)

main <- coro::async(function() {
  bars <- await(market_async$get_bars("AAPL", "1Day",
                                       start = "2024-01-01", end = "2024-01-31"))
  clock <- await(market_async$get_clock())

  print(bars)
  cat("Market open:", clock$is_open, "\n")
})

main()
while (!later::loop_empty()) later::run_now()
```

## Available Classes

| Class | Purpose |
|-------|---------|
| `AlpacaMarketData` | Historical bars, latest trades/quotes, snapshots, assets, calendar, clock, news, screener, corporate actions |
| `AlpacaTrading` | Place, modify, cancel, and query orders |
| `AlpacaAccount` | Account info, positions, portfolio history, activities, watchlists |
| `AlpacaOptions` | Options contracts, bars, trades, quotes, snapshots, chain |

Standalone function: `alpaca_backfill_bars()` -- bulk historical bar download with CSV resume.

## Citation

If you use this package in your work, please cite it:

```r
citation("alpaca")
```

> Mezquita, D. (2026). alpaca: R API Wrapper to Alpaca Trading Platform. R package version 0.1.0.

## Licence

MIT &copy; [Dereck Mezquita](https://github.com/dereckmezquita) [![ORCID](https://img.shields.io/badge/ORCID-0000--0002--9307--6762-green)](https://orcid.org/0000-0002-9307-6762). See [LICENSE.md](LICENSE.md) for the full text, including the citation clause.
