---
title: "Asynchronous Usage with alpaca"
author: "Dereck Mezquita"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Asynchronous Usage with alpaca}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  results = "hold",
  comment = "#>",
  eval = FALSE
)
```

Every R6 class in `alpaca` accepts an `async = TRUE` flag at construction.
When enabled, all methods return `promises::promise` objects instead of
direct values. This vignette shows how to consume those promises with
`coro::async`/`await` and `later::run_now`.

## Why Async?

Synchronous HTTP blocks the R session while waiting for a reply.
Asynchronous mode lets you fire off multiple requests and process results
as they arrive — useful for bots that poll several endpoints or place
orders in parallel.

## Event Loop

> **Event loop**: R does not have a built-in event loop like Node.js or
> Python's `asyncio`. Promises only resolve when the event loop ticks via
> `later::run_now()`. In scripts and vignettes, drain the loop with
> `while (!loop_empty()) run_now()`. In **Shiny** apps the event loop runs
> automatically.

## Basic Async with coro

The `coro` package provides `async()` and `await()` for writing
async code that reads like synchronous code:

```{r async-coro}
library(alpaca)
library(coro)
library(later)

md <- AlpacaMarketData$new(async = TRUE)

main <- async(function() {
  # These run concurrently
  bars_promise <- md$get_bars("AAPL", timeframe = "1Day",
                              start = "2024-01-01", end = "2024-01-31")
  clock_promise <- md$get_clock()

  bars <- await(bars_promise)
  clock <- await(clock_promise)

  cat("Got", nrow(bars), "bars for AAPL\n")
  cat("Market is", if (clock$is_open) "OPEN" else "CLOSED", "\n")
})

main()
while (!loop_empty()) run_now()
```

## Parallel Requests

Fire off multiple requests simultaneously and collect results:

```{r parallel-requests}
library(alpaca)
library(promises)
library(later)

md <- AlpacaMarketData$new(async = TRUE)
symbols <- c("AAPL", "MSFT", "GOOGL", "AMZN", "TSLA")

# Launch all requests in parallel
bar_promises <- lapply(symbols, function(sym) {
  md$get_latest_bar(sym)
})

# Collect results as they arrive
results <- list()
for (i in seq_along(bar_promises)) {
  bar_promises[[i]] %...>% (function(idx) {
    function(data) {
      results[[idx]] <<- data
    }
  })(i)
}

while (!later::loop_empty()) later::run_now()

# Combine into a single data.table
all_bars <- data.table::rbindlist(results, fill = TRUE)
print(all_bars)
```

## Async Trading

Place orders asynchronously and monitor their status:

```{r async-trading}
library(alpaca)
library(coro)
library(later)

trading <- AlpacaTrading$new(async = TRUE)
acct <- AlpacaAccount$new(async = TRUE)

main <- async(function() {
  # Place order
  order <- await(trading$add_order(
    symbol = "AAPL", side = "buy", type = "limit",
    time_in_force = "day", qty = 1, limit_price = 150
  ))
  cat("Order placed:", order$id, "\n")

  # Check account
  info <- await(acct$get_account())
  cat("Buying power:", info$buying_power, "\n")

  # Check positions
  positions <- await(acct$get_positions())
  cat("Open positions:", nrow(positions), "\n")
})

main()
while (!loop_empty()) run_now()
```

## Async in Shiny

In Shiny, the event loop runs automatically. Just return promises from
reactive expressions:

```{r shiny}
library(shiny)
library(alpaca)

md <- AlpacaMarketData$new(async = TRUE)

server <- function(input, output, session) {
  bars <- reactive({
    md$get_bars(input$symbol, timeframe = "1Day",
                start = input$start, end = input$end)
  })

  output$chart <- renderPlot({
    bars() %...>% function(data) {
      plot(data$t, data$c, type = "l",
           main = paste(input$symbol, "Close Price"),
           xlab = "Date", ylab = "Price")
    }
  })
}
```

## Error Handling

Use `promises::catch()` to handle errors in async chains:

```{r error-handling}
library(alpaca)
library(promises)
library(later)

md <- AlpacaMarketData$new(async = TRUE)

md$get_bars("INVALID_SYMBOL", timeframe = "1Day") %...>%
  function(data) {
    cat("Got", nrow(data), "bars\n")
  } %...!%
  function(err) {
    cat("Error:", conditionMessage(err), "\n")
  }

while (!later::loop_empty()) later::run_now()
```

## Next Steps

- See `vignette("getting-started")` for synchronous usage and full API
  coverage.
- Explore the [Alpaca API documentation](https://docs.alpaca.markets/) for
  endpoint details.
