---
title: "Getting Started with alpaca"
author: "Dereck Mezquita"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with alpaca}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  results = "hold",
  comment = "#>"
)

box::use(
  alpaca[
    AlpacaMarketData, AlpacaTrading, AlpacaAccount, AlpacaOptions,
    get_api_keys, get_base_url, get_data_base_url
  ]
)

box::use(../tests/testthat/mock_router[mock_router])

KEYS <- get_api_keys(
  api_key = "fake-key",
  api_secret = "fake-secret"
)
TBASE <- "https://paper-api.alpaca.markets"
DBASE <- "https://data.alpaca.markets"

options(httr2_mock = mock_router)
```

This vignette demonstrates how to use `alpaca` in **synchronous** mode to
interact with the Alpaca Trading API.

## Setup

Set your API credentials as environment variables in `.Renviron`:

```bash
ALPACA_API_KEY = your-api-key
ALPACA_API_SECRET = your-api-secret
ALPACA_API_ENDPOINT = https://paper-api.alpaca.markets
```

If you don't have a key, visit the [Alpaca dashboard](https://app.alpaca.markets/).

> **Paper trading**: Always start with
> `https://paper-api.alpaca.markets` to test without risking real money.

```{r load}
library(alpaca)

keys <- get_api_keys(
  api_key = "your-api-key",
  api_secret = "your-api-secret"
)
```

```{r keys, echo = FALSE}
keys <- KEYS
tbase <- TBASE
dbase <- DBASE
```

## Market Data

The `AlpacaMarketData` class provides access to historical and real-time
stock market data.

```{r market-init}
md <- AlpacaMarketData$new()
```

```{r market-init-hidden, echo = FALSE}
md <- AlpacaMarketData$new(keys = keys, base_url = tbase, data_base_url = dbase)
```

### Historical Bars

```{r bars}
bars <- md$get_bars(
  symbol = "AAPL",
  timeframe = "1Day",
  start = "2024-01-01",
  end = "2024-01-31"
)
bars[]
```

### Latest Trade

```{r latest-trade}
trade <- md$get_latest_trade(symbol = "AAPL")
trade[]
```

### Latest Quote

```{r latest-quote}
quote <- md$get_latest_quote(symbol = "AAPL")
quote[]
```

### Market Clock

```{r clock}
clock <- md$get_clock()
clock[]
```

### Available Assets

```{r assets}
assets <- md$get_assets(status = "active", asset_class = "us_equity")
assets[]
```

### Market Calendar

```{r calendar}
cal <- md$get_calendar(start = "2024-01-01", end = "2024-01-31")
cal[]
```

### Market News

```{r news}
news <- md$get_news(symbols = "AAPL", limit = 5)
news[, .(headline, source, created_at)]
```

---

## Account Information

The `AlpacaAccount` class gives you access to account details, positions,
and portfolio history.

```{r account-init}
acct <- AlpacaAccount$new()
```

```{r account-init-hidden, echo = FALSE}
acct <- AlpacaAccount$new(keys = keys, base_url = tbase)
```

### Account Info

```{r account-info}
info <- acct$get_account()
info[, .(status, equity, buying_power, cash)]
```

### Open Positions

```{r positions}
positions <- acct$get_positions()
positions[]
```

### Portfolio History

```{r portfolio-history}
history <- acct$get_portfolio_history(period = "1M", timeframe = "1D")
history[]
```

### Account Activities

```{r activities}
activities <- acct$get_activities(activity_types = "FILL")
activities[]
```

---

## Trading

The `AlpacaTrading` class handles order management.

```{r trading-init}
trading <- AlpacaTrading$new()
```

```{r trading-init-hidden, echo = FALSE}
trading <- AlpacaTrading$new(keys = keys, base_url = tbase)
```

### Place a Limit Order

```{r limit-order}
order <- trading$add_order(
  symbol = "AAPL",
  side = "buy",
  type = "limit",
  time_in_force = "day",
  qty = 1,
  limit_price = 150
)
order[, .(id, symbol, side, type, status)]
```

### Check Open Orders

```{r open-orders}
open_orders <- trading$get_orders(status = "open")
open_orders[]
```

### Cancel a Specific Order

```{r cancel-order}
trading$cancel_order(order_id = "order-uuid-123")
```

### Cancel All Open Orders

```{r, eval = FALSE}
# Dangerous: cancels ALL open orders
trading$cancel_all_orders()
```

---

## Options

The `AlpacaOptions` class provides access to options contracts and data.

```{r options-init}
opts <- AlpacaOptions$new()
```

```{r options-init-hidden, echo = FALSE}
opts <- AlpacaOptions$new(keys = keys, base_url = tbase, data_base_url = dbase)
```

### Search Contracts

```{r contracts}
contracts <- opts$get_contracts(
  underlying_symbols = "AAPL",
  type = "call",
  expiration_date_gte = "2024-06-01",
  limit = 10
)
contracts[, .(symbol, type, strike_price, expiration_date)]
```

### Options Chain

```{r option-chain}
chain <- opts$get_option_chain(underlying_symbol = "AAPL", type = "call")
chain[]
```

---

## Position Management

```{r, eval = FALSE}
# Close 50% of a position
acct$close_position(symbol_or_id = "AAPL", percentage = 50)

# Close entire position
acct$close_position(symbol_or_id = "AAPL")

# Close all positions (dangerous!)
acct$close_all_positions(cancel_orders = TRUE)
```

## Next Steps

- See `vignette("async-usage")` for promise-based async workflows.
- See `vignette("margin-short-selling")` for margin and short selling.
- Explore the [Alpaca API documentation](https://docs.alpaca.markets/) for
  full endpoint details.
