---
title: "Getting Started with alpaca"
author: "Dereck Mezquita"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with alpaca}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette demonstrates how to use `alpaca` in **synchronous** mode to
interact with the Alpaca Trading API.

## Setup

Set your API credentials as environment variables in `.Renviron`:

```bash
ALPACA_API_KEY = your-api-key
ALPACA_API_SECRET = your-api-secret
ALPACA_API_ENDPOINT = https://paper-api.alpaca.markets
```

If you don't have a key, visit the [Alpaca dashboard](https://app.alpaca.markets/).

> **Paper trading**: Always start with
> `https://paper-api.alpaca.markets` to test without risking real money.

## Market Data

The `AlpacaMarketData` class provides access to historical and real-time
stock market data.

```{r market-data}
library(alpaca)

md <- AlpacaMarketData$new()

# Historical daily bars for AAPL
bars <- md$get_bars("AAPL", timeframe = "1Day", start = "2024-01-01",
                    end = "2024-01-31")
print(bars)

# Latest trade
latest <- md$get_latest_trade("AAPL")
print(latest)

# Market clock
clock <- md$get_clock()
print(clock)

# Available assets
assets <- md$get_assets(status = "active", asset_class = "us_equity")
print(head(assets))
```

## Account Information

The `AlpacaAccount` class gives you access to account details, positions,
and portfolio history.

```{r account}
acct <- AlpacaAccount$new()

# Account info
info <- acct$get_account()
print(info[, .(status, equity, buying_power, cash)])

# Open positions
positions <- acct$get_positions()
print(positions)

# Portfolio history
history <- acct$get_portfolio_history(period = "1M", timeframe = "1D")
print(history)
```

## Trading

The `AlpacaTrading` class handles order management.

```{r trading}
trading <- AlpacaTrading$new()

# Place a limit order (paper trading)
order <- trading$add_order(
  symbol = "AAPL",
  side = "buy",
  type = "limit",
  time_in_force = "day",
  qty = 1,
  limit_price = 150
)
print(order[, .(id, symbol, side, type, status)])

# Check open orders
open_orders <- trading$get_orders(status = "open")
print(open_orders)

# Cancel a specific order
trading$cancel_order(order$id)

# Cancel all open orders
trading$cancel_all_orders()
```

### Order Types

`alpaca` supports all Alpaca order types:

```{r order-types}
# Market order by dollar amount
trading$add_order(
  symbol = "AAPL", side = "buy", type = "market",
  time_in_force = "day", notional = 500
)

# Stop-limit order
trading$add_order(
  symbol = "AAPL", side = "sell", type = "stop_limit",
  time_in_force = "gtc", qty = 10,
  stop_price = 140, limit_price = 139
)

# Trailing stop order
trading$add_order(
  symbol = "AAPL", side = "sell", type = "trailing_stop",
  time_in_force = "gtc", qty = 10,
  trail_percent = 5
)

# Bracket order
trading$add_order(
  symbol = "AAPL", side = "buy", type = "market",
  time_in_force = "day", qty = 10,
  order_class = "bracket",
  take_profit = list(limit_price = 200),
  stop_loss = list(stop_price = 140, limit_price = 139)
)
```

## Options

The `AlpacaOptions` class provides access to options contracts and data.

```{r options}
opts <- AlpacaOptions$new()

# Find AAPL call options expiring after 2024-06-01
contracts <- opts$get_contracts(
  underlying_symbols = "AAPL",
  type = "call",
  expiration_date_gte = "2024-06-01",
  limit = 10
)
print(contracts[, .(symbol, type, strike_price, expiration_date)])
```

## Position Management

```{r positions}
acct <- AlpacaAccount$new()

# View a specific position
pos <- acct$get_position("AAPL")
print(pos[, .(symbol, qty, avg_entry_price, current_price, unrealized_pl)])

# Close 50% of a position
acct$close_position("AAPL", percentage = 50)

# Close entire position
acct$close_position("AAPL")
```

## Next Steps

- See `vignette("async-usage")` for promise-based async workflows.
- Explore the [Alpaca API documentation](https://docs.alpaca.markets/) for
  full endpoint details.
