---
title: "Margin Trading and Short Selling with alpaca"
author: "Dereck Mezquita"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Margin Trading and Short Selling with alpaca}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette covers margin trading, short selling, and options position
management using the `alpaca` package. Unlike crypto exchanges which require
separate margin/futures classes, Alpaca integrates margin and short selling
directly into the standard trading API.

## How Margin Works on Alpaca

Alpaca accounts support margin trading out of the box:

- **Reg-T margin**: 2x buying power for overnight positions
- **Day-trading margin**: 4x buying power for positions opened and closed
  same-day (requires PDT flag and $25k+ equity)
- **Short selling**: Sell stocks you don't own; Alpaca borrows shares
  automatically

No special endpoints or classes are needed — `AlpacaTrading` and
`AlpacaAccount` handle everything.

## Setup

```{r load}
library(alpaca)

trading <- AlpacaTrading$new()
acct <- AlpacaAccount$new()
market <- AlpacaMarketData$new()
```

## Check Margin Status

```{r margin-status}
info <- acct$get_account()

# Key margin fields
info[, .(
  equity,
  buying_power,         # Total buying power (includes margin)
  regt_buying_power,    # Reg-T (2x) buying power
  daytrading_buying_power, # 4x for day trades (if PDT)
  initial_margin,       # Current initial margin requirement
  maintenance_margin,   # Maintenance margin requirement
  long_market_value,    # Total long positions
  short_market_value,   # Total short positions
  multiplier,           # Account leverage multiplier (2 or 4)
  pattern_day_trader    # PDT flag
)]
```

## Check if a Stock is Shortable/Marginable

```{r check-shortable}
# Check asset attributes
asset <- market$get_asset("AAPL")
cat("Shortable:", asset$shortable, "\n")
cat("Marginable:", asset$marginable, "\n")
cat("Fractionable:", asset$fractionable, "\n")

# Filter for all shortable stocks
assets <- market$get_assets(status = "active", asset_class = "us_equity")
shortable <- assets[shortable == TRUE]
cat("Shortable stocks:", nrow(shortable), "\n")
```

## Short Selling

Short selling on Alpaca is simply placing a **sell order** for a stock you
don't currently own. Alpaca handles the borrow automatically.

### Open a Short Position

```{r short-open}
# Short 100 shares of AAPL at market
order <- trading$add_order(
  symbol = "AAPL",
  side = "sell",
  type = "market",
  time_in_force = "day",
  qty = 100
)
print(order[, .(id, symbol, side, type, status)])
```

### Monitor the Short Position

```{r short-monitor}
# Check the position — side will be "short"
pos <- acct$get_position("AAPL")
print(pos[, .(
  symbol, qty, side,
  avg_entry_price, current_price,
  unrealized_pl, unrealized_plpc
)])
```

### Close the Short (Buy to Cover)

```{r short-close}
# Option 1: Buy the shares back manually
trading$add_order(
  symbol = "AAPL",
  side = "buy",
  type = "market",
  time_in_force = "day",
  qty = 100
)

# Option 2: Use close_position (simpler)
acct$close_position("AAPL")

# Option 3: Close 50% of the short
acct$close_position("AAPL", percentage = 50)
```

### Short with a Bracket Order (Risk Management)

```{r short-bracket}
# Short with automatic take-profit and stop-loss
order <- trading$add_order(
  symbol = "TSLA",
  side = "sell",
  type = "limit",
  time_in_force = "gtc",
  qty = 50,
  limit_price = 250,
  order_class = "bracket",
  take_profit = list(limit_price = 200),   # Buy back at $200 for profit

  stop_loss = list(stop_price = 275, limit_price = 280)  # Stop loss at $275
)
```

## Margin Long Positions

With a margin account, you can buy more shares than your cash allows.

```{r margin-long}
# Your buying power is 2x (or 4x for day trades) your equity
info <- acct$get_account()
cat("Cash:", info$cash, "\n")
cat("Buying power:", info$buying_power, "\n")

# Buy $200k worth of stock with only $100k cash
order <- trading$add_order(
  symbol = "SPY",
  side = "buy",
  type = "market",
  time_in_force = "day",
  notional = 200000  # Dollar-based order
)
```

## Options Trading with Position Intent

For options, use `position_intent` to signal your intent:

```{r options-intent}
opts <- AlpacaOptions$new()

# Find an options contract
contracts <- opts$get_contracts(
  underlying_symbols = "AAPL",
  type = "call",
  expiration_date_gte = "2024-06-01",
  limit = 5
)
print(contracts[, .(symbol, type, strike_price, expiration_date)])

# Buy to open a call option
order <- trading$add_order(
  symbol = contracts$symbol[1],
  side = "buy",
  type = "limit",
  time_in_force = "day",
  qty = 1,
  limit_price = 5.00,
  position_intent = "buy_to_open"
)

# Sell to close the option
trading$add_order(
  symbol = contracts$symbol[1],
  side = "sell",
  type = "limit",
  time_in_force = "day",
  qty = 1,
  limit_price = 7.50,
  position_intent = "sell_to_close"
)
```

## Corporate Actions

Monitor dividends and splits that affect your positions:

```{r corporate-actions}
# Get upcoming dividends for your positions
divs <- market$get_corporate_actions(
  ca_types = "dividend",
  since = "2024-01-01",
  until = "2024-12-31",
  symbol = "AAPL"
)
print(divs[, .(ca_type, initiating_symbol, ex_date, cash)])

# Check for stock splits
splits <- market$get_corporate_actions(
  ca_types = "split",
  since = "2024-01-01",
  until = "2024-12-31"
)
```

## News-Driven Trading

Use the news endpoint for event-driven strategies:
```{r news}
# Get latest news for your watchlist
news <- market$get_news(symbols = "AAPL,TSLA,NVDA", limit = 10)
print(news[, .(headline, source, created_at)])
```

## Portfolio Risk Monitoring

```{r risk}
# Full position overview
positions <- acct$get_positions()
if (nrow(positions) > 0) {
  positions[, .(
    symbol, side, qty,
    market_value, unrealized_pl, unrealized_plpc,
    change_today
  )]
}

# Portfolio history for P/L tracking
history <- acct$get_portfolio_history(period = "1M", timeframe = "1D")
print(history)

# Account activities (fills, dividends, etc.)
fills <- acct$get_activities_by_type("FILL")
print(fills)
```

## Emergency: Close Everything

```{r emergency}
# Cancel all open orders and close all positions
acct$close_all_positions(cancel_orders = TRUE)
```

## Next Steps

- See `vignette("getting-started")` for basic market data and order placement.
- See `vignette("async-usage")` for promise-based async workflows.
- Explore the [Alpaca API documentation](https://docs.alpaca.markets/) for
  full endpoint details.
